Hardware-accelerated Rendering of Intersecting Volumes using Boolean Operations

### 2013-05-26 paper.tex
Added new screenshots (ref #9)
Added basic table with some performance measurements (ref #4)
Frame rate for slicing on Macbook (ref #4)
Frame rate for optimized slicing on Macbook (ref #4)
Removed bullets in Evaluation section (ref #14)
Initial evaluation section (ref #14)

\documentclass{article}
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{fancyvrb}

@@

\begin{figure}
\centering
\includegraphics[width=0.95\textwidth]{boolean-and-screenshot-1.png}
\caption{Screenshot of Boolean AND.}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.95\textwidth]{boolean-xor-screenshot-1.png}
\caption{Screenshot of Boolean XOR.}
\end{figure}

@@

\begin{itemize}
  \item How fast can they be rendered?
  \item How easy were they to implement?
\end{itemize}

>>>>

\documentclass{article}
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{fancyvrb}

@@

\begin{figure}
\centering
\includegraphics[width=0.95\textwidth]{boolean-and-screenshot-2.png}
\caption{Screenshot of Boolean AND.}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.95\textwidth]{boolean-xor-screenshot-2.png}
\caption{Screenshot of Boolean XOR.}
\end{figure}

@@

The two approaches to rendering intersecting volumes using boolean operations
were evaluated using different criteria.

Because volume visualization is used in the scientific and medical communities
to aid in making important decisions, an approach that does not handle all cases
will likely not be appropriate for widespread usage.  With that in mind, the
accuracy of each approach was judged by looking for edge cases with visual
artifacts.

Unfortunately the Boolean XOR approach may show some visual artifacts from
certain angles.  This occurs because when two pieces have the same depth value,
due to perspective one of the pieces may actually appear behind the other.  In
the future, this issue may be solved by using the distance from the camera to
sort rather than the actual depth.

% Picture of visual artifacts with Boolean XOR approach

% Picture of depth-sorting issue

The performance of each approach was measured by using a simple frame counter
showing the number of frames rendered in the previous second.  Multiple camera
angles and scene configurations were tried in order to expose strong and weak
points of the approaches.  The tests were run on at least three different
machines with a variety of operating systems (Linux, Mac OS X) and video cards
(ATI, NVIDIA).  Table \ref{performance-table} shows the results.

\begin{table}
  \centering
  \begin{tabular}{ l l l }
    \toprule
    Approach & Macbook & Desktop \\
    \midrule
    Boolean AND with masking & 22 & 13 \\
    Boolean AND with XOR & 24 & 13 \\
    Slicing & 6 & ? \\
    Optimized Slicing & 25 & ? \\
    \bottomrule
  \end{tabular}
  \caption{Performance measurements}
  \label{performance-table}
\end{table}

% Furthermore, a variety of sampling rates were used to determine how each
% approach scales as the number of samples taken increases.

As a baseline the Boolean-based approaches were compared to the existing
view-aligned slicing method from previous work by developing a simple prototype.
Since the slicing method requires significantly more state changes when the
volumes are side-by-side as opposed to when one is in front of the other, it was
particularly important with this method that performance measurements were taken
from a variety of perspectives.

In addition, a variation of the slicing method that was optimized for two
volumes was created.  Basically rather than using uniform variables to switch
textures as would normally be done, the optimized variation stores the texture
with each vertex instead.  While it requires a fairly significant amount of
extra storage, using vertex attributes allows all the slices to be transferred
to the card at once.  The result is a considerable performance boost, similar to
how it is generally faster to transfer one large file across a network then many
small ones.

As shown both Boolean-based approaches outperformed the traditional slicing
method, but were in turn bested by the optimized slicing method.  It should be
noted that the frame rates for the optimized slicing method are the worst case,
and in average it performed even better, around forty frames per second.

Judging from the numbers, it appears that the multiple passes of the
Boolean-based approaches create too much overhead compared to the optimized
slicing method.  The optimized slicing method also has a very simple fragment
shader which may keep it from being fragment-bound.

Altogether, although the Boolean-based approaches outperform the un-optimized
slicing method, it is recommended that the traditional slicing method with the
optimizations described be used.  The optimized slicing method is not only
faster, but it supports volumes of any orientation.  Furthermore, the optimized
method should scale to more volumes relatively easily.
